\documentclass[a4paper, 11pt]{article}
\usepackage{comment} % enables the use of multi-line comments (\ifx \fi)  
\usepackage{fullpage} % changes the margin
\usepackage{mathtools} %allows us to write complex equations
\usepackage{graphicx} %allows us to add pictures
\usepackage{amsmath} %allows us to add Greek letters and equations
\usepackage{amssymb}
\usepackage{float} %formatting pictures
\usepackage{afterpage}
\usepackage{tikz}
\usepackage{minted}
\usepackage{filecontents,listings,varwidth}
\usetikzlibrary{automata, positioning, arrows}
\tikzset{->,>=stealth',node distance=3cm}

\begin{document}
\noindent
\large\textbf{Final Exam} \hfill \\
\large{Mike Gao} \\
\normalsize 260915701 \\
Prof. Prakash Panangaden \hfill 


\section*{Question 1}
$\epsilon + (b^*ba)^*b^*b$

\section*{Question 2}
Minimize the dfa and check if it only has one state. If yes, check whether the state is accepting, if yes, then it accepts $\Sigma^*$

\section*{Question 3}
\subsection*{3.1}
Decidable. Since the intersection of a context free language and a regular language is context free, and we can build context-free grammar from the language. For a given CFG $G$, checking whether L(G) is empty or not is decidable.
\subsection*{3.2}
It is a well known problem that $L=\Sigma^*$ is undecidable. We combine the PDA of L and DFA of R to create a new PDA. Because of the CFL-PDA equivalence, it is undecidable whether $L \cap R = \Sigma^*$.
\section*{Question 4}
Context free but not regular.

\noindent The following grammar generates it:

$S \rightarrow aSc | T | \epsilon$  $T \rightarrow bTc | \epsilon$

\noindent Thus it is context free. Now we need to prove that it is not regular:

\noindent Regular languages are closed under intersection. We intersect L with another regular language, we get $L \cap \{b^*c^*\} = \{b^mc^m\} $ which is not regular. According to the closure property of regular language, we get $L$ is not regular language as well.

\section*{Question 5}
\subsection*{5.1}
Suppose you have an input on which M does not halt. This detection algorithm will run forever. Therefore it is undecidable.
\subsection*{5.2}
Consider the following set $\overline {H_{TM}}=\{<M,w> | $M does not halt on w$\}$ Such set is non-CE as we mentioned in class. 
We do a reduction $\overline {H_{TM}} \leq_m  FIN$

Construct a new turing machine $M'$ with input x, and let l be length of x. 

We run w on M for $l$ steps.

There are two cases: $(\overline{H_{TM}})$ If M does not halt during the $l$ steps, nothing is accepted. So $M'=nil\in FIN$

$(H_{TM})$ If M halts, all strings x with length $\geq l$ will be accepted by $M'$, which means $M'$ accepts infinitely many strings, so $M' \notin FIN$

We know that $\overline {H_{TM}}$ is non-CE, then $FIN$ is non-CE as well.


\section*{Question 6}
\subsection*{6.1}
Decidable. Whether a regular language is co-finite is decidable. The regular languages are closed under complement so this is essentially asking whether regular language is finite, which is decidable.
\subsection*{6.2}
Not decidable. We know that a set of invalid computations is a CFL. Such set is co-finite iff the Turing machine can only accept a finite set. Since it is undecidable to determine whether a given Turing machine accepts a finite or infinite number of inputs (Corollary from Rice's theorem), the statement is undecidable as well.

\subsection*{6.3}
Not decidable by Rice's Theorem.

\section*{Question 7}
\subsection*{7.1}
True
\subsection*{7.2}
True
\subsection*{7.3}
True
\subsection*{7.4}
True
\subsection*{7.5}
False
\newpage
\section*{Honor Code}
\noindent I solemnly swear that I am up to no mischief. I did not consult
anyone nor did I use the internet to search for answers to these
questions

\begin{filecontents*}{prakash.txt}
##################################################################
##################################################################
###########################%%%%##########((#(((###################
######################%%%%%#(((////////(((###(((##################
###################%%&%%%#((///////////////((((#((((((############
####(((((#(########%&&%##(//////**********///(((((((((####((((####
#(((((((((((((((#%%%&&%##(//***/************///((/(((((((((((((#(#
(((((((((((((((#%%%%%&%##(//******,,,,,,,***/////(/(((((((((((((((
(((((((((((((((##%%%%%%##(//******,,,,,,,***/////(((((((((((((((((
(((((((((((((((%%%%%%%%#((/*****,,,,,,,,,**////////(((((((((((((((
(((((((((((((%%%&&&%%%%##((((/***,,*,*****////////(/((((((((((((((
(((((((((((((####%&&%%&%#(//###(/***//(/***//((/%%/*//((((((((((((
((((((((((((/####&%%(#%%##%(((#%#**#(//(&&*###(##/****((((((((((((
(((((((((((//(%##&&%#((/(/////##(***#*****/////((****/((((((((((((
((((((((((////(##%&%%#(//***/%%(/***/*,/((//***(/*/*//((((((((((((
((((((((////////##&&%#(//*(#%##(/******/(*,,**//**//////((((((((((
((((//////////////#&%##((((#%###((//**/**////////*/////////(((((((
((/////////////////&%%#(/(#%&%#((////((/*/*////***////////////((((
///////////////////@&%%%(#(((#(((((//////////*********////////((((
/////////////////@@@%%&%%%#(((//*****////((#************///////(((
/////////////**#@@@@@%%%&&&%#(//((////(%((&@&#***********/////////
/////////*(&@@@@@@@@@@%%%%%%%%%%%###((((/(@@@&%**********/////////
/////%&%@@@@@@@@@@@@@@@%###(#%#((////((///@%&&@&&&%%%/****////////
(&&#%&&&&%&%&@@&@@@@@@@@&#(//((((////(////@&@@@&@%%&&&%%%%%///////
&&&&&&%#%&@&%@@&&&&&@@&@&&&#(/////////#**#&%&@@@&@&&&&&&&@&&&&#%//
@@&&&@@%&@&&&@@@&@&%&@@%@@&@&/**********(&&&&&&@%&&&%&&&&@&&&&&#%&
&@&&&&@@%&@&@@@@@&&#&&&@@@@@@@@%(******(&&&&%&@@&#%&%&%%&@@&@&&#&&
@@@&&&@@&%@@&%@@&&%%%%&&&&#%%&(@@@@(,*(%&%%@%%%%%&@&%&&&&@&%&&&#&&
\end{filecontents*}

\noindent\lstinputlisting[basicstyle=\ttfamily]{prakash.txt}

\end{document}
